Cycles of the logistic map

### 2013-05-03 doc/pc-ijbc.tex
edit introduction
improve the webpage

\begin{abstract}
%\abstract{
%
The onset and bifurcation points of the $n$-cycles of
  a polynomial map are located
  through a characteristic equation
  connecting the cyclic polynomials of the periodic orbit points.
The minimal polynomials of the critical parameters
  of the logistic, H\'enon, and cubic maps are obtained
  for $n$ up to 13, 9, and 8,
  respectively.
%
%}
\end{abstract}

@@

A fixed point is a solution of $x^* = f(x^*)$.
%
If $x_1$ deviates slightly from $x^*$
  and the sequence still converges to $x^*$,
  then it is \emph{stable}.
%
For a differentiable $f$,
  a stable fixed point requires $|f'(x^*)| \le 1$
  to reduce deviations in successive iterations \cite{strogatz}.

In an $n$-cycle, $n$ is the smallest positive integer
  that allows $x_1 = x_{n+1} = f^n(x_1)$,
  where $f^n$ is the $n$th iterate of $f$,
  e.g., $f^3(x) = f(f(f(x)))$.
Thus, any $x_k$ in an $n$-cycle of $f$ must be a fixed point of $f^n$
%
\big[the converse is, however, untrue, for a fixed point of $f^n$
  can also be a fixed point of $f^d$ for a divisor $d$ of $n$:
  if $f^d(x) = x$, then $f^n(x) = f^d(\cdots f^d(x)\cdots) = x$\big],
%
%
and a cycle can be classified as stable or unstable
  according to the corresponding fixed point of $f^n$:
  a stable cycle requires
  $\big| \frac {d} {dx} f^n(x_1) \big| \le 1$,
  or by the chain rule,
%
%
%
\begin{equation}
  \Big| f'(x_1) \cdots f'(x_n) \Big| \le 1,
\label{eq:der}
\end{equation}
The set of the $n$ points within the cycle $x_1. \ldots, x_n$
   is called the \emph{orbit}.
The onset and bifurcation points
  are defined at the loci
  where $\frac {d} {dx} f^n(x_1)$ reaches $+1$ and $-1$,
  respectively \cite{strogatz}.
%
%
%
The orbit and cycle stability depend on the parameter $r$.
%
Below we will present an algorithm to identify
  all regions of $r$ that allow stable $n$-cycles.

@@

To summarize, we have
\begin{theorem}
  At the onset and bifurcation points,
  the square-free cyclic polynomials $C_p(\vx)$
  of the orbit points $\vx = \{x_k\}$
  are linear related by % \refeq{xcp},
  $\lambda \, C_p(\vx) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx)$,
  with $\lambda$ being $+1$ and $-1$, respectively,
  and $T_{pq}(R)$ being the coefficients
  from the square-free reduction of the cyclic polynomial
  $C_p(\vx) \, \prod_k f'(x_k)$. %[\refeq{logder}].
  Thus, the fixed point of $f^n$
  is marginally stable when $R$ satisfies
  characteristic equation $\big| \lambda \, \vct I - \vct T(R) \big| = 0$.
  \label{thm:main}
\end{theorem}

@@

\begin{theorem}
The number $\NB(n)$ of the square-free cyclic polynomials $C_p(\vx)$
  formed by the orbit points $\vx = \{x_1, \ldots, x_n\}$,
  which is also $\deg_\lambda A_n\Rlam$, is given by
\begin{equation}
  %\deg_\lambda A_n\Rlam = 
  \NB(n) = N(n) = (1/n) \textstyle\sum_{d|n} \varphi(n/d) 2^{\,d}.
\label{eq:necklace}
\end{equation}
%
\label{thm:necklace}
\end{theorem}
%
Here, $\varphi(m)$ is Euler's totient function,
which gives the number of integers from 1 to $m$
  that are coprime to $m$,
  e.g.,
  $\varphi(1) = 1$,
  $\varphi(3) = 2$ for 1 and 2,
  $\varphi(6) = 2$ for 1 and 5.
The first few values of $N_\Pset(n)$ are 2, 3, 4, 6, 8, 14, 20, 36, 60, 108, 188, 352, 632, \ldots,
  starting from $n = 1$.

@@

\begin{theorem}
The degree in $\lambda$ of the minimal polynomial $P_n\Rlam$ of all $n$-cycles
  is equal to the number of the $n$-cycles,
  which is given by
\begin{equation}
  %\deg_\lambda P_n\Rlam = 
  L(n) = (1/n)\textstyle\sum_{d | n} \mu(n/d) 2^{\,d}.
\label{eq:lyndon}
\end{equation}
\label{thm:lyndon}
\end{theorem}
%
%
\begin{proof}
Let $\lambda_c$ be the value of $\Lambda(\vx^c)$
  evaluated at the $c$th orbit points $\vx^c$.
The minimal polynomial $P_n\Rlam$
  can be written as $\prod_c (\lambda - \lambda_c)$,
  and $\deg_\lambda P_n\Rlam$ is the same as
  the number $L(n)$ of $n$-cycles, which is known \cite{hao, hao2, lutzky}.
\end{proof}
%
%
The first few values of $L(n)$ are
2, 1, 2, 3, 6, 9, 18, 30, 56, 99, 186, 335, 630, \ldots,
from $n = 1$.
%
Incidentally, $L(n)$ is also the number of aperiodic necklaces of length $n$.
  In the $n = 4$ case,
  $0001$, $0011$, and $0111$ are aperiodic;
  but $0000$, $1111$, and $0101$ are not
  owing to the shorter periods 1, 2, and 2, respectively.
%
Since an $n$-necklace is either aperiodic or
  a repeated aperiodic $d$-necklace (with $d|n$),
  we have
\begin{equation}
  N(n) = \textstyle\sum_{d|n} L(d).
\label{eq:necklacelyndon}
\end{equation}
%

>>>>

\begin{abstract}
%\abstract{
%
The onset and bifurcation points of the $n$-cycles of
  a polynomial map are located
  through a characteristic equation
  connecting the cyclic polynomials of the cycle points.
The minimal polynomials of the critical parameters
  of the logistic, H\'enon, and cubic maps are obtained
  for $n$ up to 13, 9, and 8,
  respectively.
%
%}
\end{abstract}

@@

A fixed point is a solution of $x^* = f(x^*)$.
%
It is stable if $x_1$ deviates slightly from $x^*$
  and the sequence still converges to $x^*$.
%
For a differentiable $f$,
  a stable fixed point requires $|f'(x^*)| \le 1$
  to reduce deviations in successive iterations \cite{strogatz}.

An $n$-cycle is a sequence $x_1, \dots, x_n$,
  such that $n$ is the smallest positive integer
  that satisfies $x_1 = x_{n+1} = f^n(x_1)$,
  where $f^n$ is the $n$th iterate of $f$,
  e.g., $f^3(x) = f(f(f(x)))$.
Obviously, any $x_k$ in an $n$-cycle of $f$ is a fixed point of $f^n$
%
\big[but the converse is untrue, for a fixed point of $f^n$
  can also be a fixed point of $f^d$ for a divisor $d$ of $n$:
  if $f^d(x) = x$, then $f^n(x) = f^d(\cdots f^d(x)\cdots) = x$\big].
%
%
Thus, a cycle can be classified as stable or unstable
  according to the corresponding fixed point of $f^n$:
  a stable cycle requires
  $\big| \frac {d} {dx} f^n(x_1) \big| \le 1$,
  or by the chain rule,
%
%
%
\begin{equation}
  \Big| f'(x_1) \cdots f'(x_n) \Big| \le 1,
\label{eq:der}
\end{equation}
%
%
The onset and bifurcation points
  are defined at the loci
  where $\frac {d} {dx} f^n(x_1)$ reaches $+1$ and $-1$,
  respectively \cite{strogatz}.
%
%
%
The cycle and its stability depend on the parameter $r$.
%
Our purpose is to present an algorithm to identify
  all regions of $r$ that allow stable $n$-cycles.

@@

To summarize, we have
\begin{theorem}
  At the onset and bifurcation points,
  the square-free cyclic polynomials $C_p(\vx)$
  of the cycle points $\vx = \{x_k\}$
  are linear related by % \refeq{xcp},
  $\lambda \, C_p(\vx) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx)$,
  with $\lambda$ being $+1$ and $-1$, respectively,
  and $T_{pq}(R)$ being the coefficients
  from the square-free reduction of the cyclic polynomial
  $C_p(\vx) \, \prod_k f'(x_k)$. %[\refeq{logder}].
  Thus, the fixed point of $f^n$
  is marginally stable when $R$ satisfies
  characteristic equation $\big| \lambda \, \vct I - \vct T(R) \big| = 0$.
  \label{thm:main}
\end{theorem}

@@

\begin{theorem}
The number $\NB(n)$ of the square-free cyclic polynomials $C_p(\vx)$
  formed by the cycle points $\vx = \{x_1, \ldots, x_n\}$,
  which is also $\deg_\lambda A_n\Rlam$, is given by
\begin{equation}
  %\deg_\lambda A_n\Rlam =
  \NB(n) = N(n) = (1/n) \textstyle\sum_{d|n} \varphi(n/d) 2^{\,d}.
\label{eq:necklace}
\end{equation}
%
\label{thm:necklace}
\end{theorem}
%
Here, $\varphi(m)$ is Euler's totient function,
which gives the number of integers from 1 to $m$
  that are coprime to $m$,
  e.g.,
  $\varphi(1) = 1$,
  $\varphi(3) = 2$ for 1 and 2,
  $\varphi(6) = 2$ for 1 and 5.
The first few values of $N_\Pset(n)$ are 2, 3, 4, 6, 8, 14, 20, 36, 60, 108, 188, 352, 632, \ldots,
  starting from $n = 1$.

@@

\begin{theorem}
The degree in $\lambda$ of the minimal polynomial $P_n\Rlam$ of all $n$-cycles
  is equal to the number of the $n$-cycles,
  which is given by
\begin{equation}
  %\deg_\lambda P_n\Rlam =
  L(n) = (1/n)\textstyle\sum_{d | n} \mu(n/d) 2^{\,d}.
\label{eq:lyndon}
\end{equation}
\label{thm:lyndon}
\end{theorem}
%
%
\begin{proof}
Let $\lambda_c$ be the value of $\Lambda(\vx^c)$
  evaluated at the $c$th cycle.
The minimal polynomial $P_n\Rlam$
  can be written as $\prod_c (\lambda - \lambda_c)$,
  and $\deg_\lambda P_n\Rlam$ is the same as
  the number $L(n)$ of $n$-cycles, which is known \cite{hao, hao2, lutzky}.
\end{proof}
%
%
The first few values of $L(n)$ are
2, 1, 2, 3, 6, 9, 18, 30, 56, 99, 186, 335, 630, \ldots,
from $n = 1$.
%
Incidentally, $L(n)$ is also the number of aperiodic necklaces of length $n$.
  In the $n = 4$ case,
  $0001$, $0011$, and $0111$ are aperiodic;
  but $0000$, $1111$, and $0101$ are not
  owing to the shorter periods 1, 2, and 2, respectively.
%
Since an $n$-necklace is either aperiodic or
  a repeated aperiodic $d$-necklace (with $d|n$),
  we have
\begin{equation}
  N(n) = \textstyle\sum_{d|n} L(d).
\label{eq:necklacelyndon}
\end{equation}
%

### 2013-05-03 doc/pc.tex
edit introduction
improve the webpage

\begin{abstract}
%\abstract{
%
The onset and bifurcation points of the $n$-cycles of
  a polynomial map are located
  through a characteristic equation
  connecting the cyclic polynomials of the periodic orbit points.
The minimal polynomials of the critical parameters
  of the logistic, H\'enon, and cubic maps are obtained
  for $n$ up to 13, 9, and 8,
  respectively.
%
%}
\end{abstract}
\maketitle

@@

A fixed point is a solution of $x^* = f(x^*)$.
%
If $x_1$ deviates slightly from $x^*$
  and the sequence still converges to $x^*$,
  then it is \emph{stable}.
%
For a differentiable $f$,
  a stable fixed point requires $|f'(x^*)| \le 1$
  to reduce deviations in successive iterations \cite{strogatz}.

In an $n$-cycle, $n$ is the smallest positive integer
  that allows $x_1 = x_{n+1} = f^n(x_1)$,
  where $f^n$ is the $n$th iterate of $f$,
  e.g., $f^3(x) = f(f(f(x)))$.
Thus, any $x_k$ in an $n$-cycle of $f$ must be a fixed point of $f^n$
%
\big[the converse is, however, untrue, for a fixed point of $f^n$
  can also be a fixed point of $f^d$ for a divisor $d$ of $n$:
  if $f^d(x) = x$, then $f^n(x) = f^d(\cdots f^d(x)\cdots) = x$\big],
%
%
and a cycle can be classified as stable or unstable
  according to the corresponding fixed point of $f^n$:
  a stable cycle requires
  $\big| \frac {d} {dx} f^n(x_1) \big| \le 1$,
  or by the chain rule,
%
%
%
\begin{equation}
  \Big| f'(x_1) \cdots f'(x_n) \Big| \le 1,
\label{eq:der}
\end{equation}
The set of the $n$ points within the cycle $x_1. \ldots, x_n$
   is called the \emph{orbit}.
The onset and bifurcation points
  are defined at the loci
  where $\frac {d} {dx} f^n(x_1)$ reaches $+1$ and $-1$,
  respectively \cite{strogatz}.
%
%
%

The orbit and cycle stability depend on the parameter $r$.
%
Below we will present an algorithm to identify
  all regions of $r$ that allow stable $n$-cycles.

>>>>

\begin{abstract}
%\abstract{
%
The onset and bifurcation points of the $n$-cycles of
  a polynomial map are located
  through a characteristic equation
  connecting the cyclic polynomials of the cycle points.
The minimal polynomials of the critical parameters
  of the logistic, H\'enon, and cubic maps are obtained
  for $n$ up to 13, 9, and 8,
  respectively.
%
%}
\end{abstract}
\maketitle

@@

A fixed point is a solution of $x^* = f(x^*)$.
%
It is stable if $x_1$ deviates slightly from $x^*$
  and the sequence still converges to $x^*$.
%
For a differentiable $f$,
  a stable fixed point requires $|f'(x^*)| \le 1$
  to reduce deviations in successive iterations \cite{strogatz}.

An $n$-cycle is a sequence $x_1, \dots, x_n$,
  such that $n$ is the smallest positive integer
  that satisfies $x_1 = x_{n+1} = f^n(x_1)$,
  where $f^n$ is the $n$th iterate of $f$,
  e.g., $f^3(x) = f(f(f(x)))$.
Obviously, any $x_k$ in an $n$-cycle of $f$ is a fixed point of $f^n$
%
\big[but the converse is untrue, for a fixed point of $f^n$
  can also be a fixed point of $f^d$ for a divisor $d$ of $n$:
  if $f^d(x) = x$, then $f^n(x) = f^d(\cdots f^d(x)\cdots) = x$\big].
%
%
Thus, a cycle can be classified as stable or unstable
  according to the corresponding fixed point of $f^n$:
  a stable cycle requires
  $\big| \frac {d} {dx} f^n(x_1) \big| \le 1$,
  or by the chain rule,
%
%
%
\begin{equation}
  \Big| f'(x_1) \cdots f'(x_n) \Big| \le 1,
\label{eq:der}
\end{equation}
%
%
%The set of the $n$ points within the cycle $x_1. \ldots, x_n$
%   is called the \emph{orbit}.
The onset and bifurcation points
  are defined at the loci
  where $\frac {d} {dx} f^n(x_1)$ reaches $+1$ and $-1$,
  respectively \cite{strogatz}.
%
%
%

The orbit and cycle stability depend on the parameter $r$.
%
Our purpose is to present an algorithm to identify
  all regions of $r$ that allow stable $n$-cycles.
